---
title: "Getting geographic metadata from UK postcodes"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{postcodes-vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

I now show a standard way to get geographic metadata on UK postcodes, using the `postcodes` and `postcodes_metadata` functions in this package.

<br/>

Imagine that we want to know more about where six people live, but we only know their `id` and something about their `postcode`.

```{r, warning=FALSE, message=FALSE}
library(DataKindR)
library(dplyr)
library(purrr)
library(tibble)
library(tidyr)

data <- tibble(
  id = c("A1", "B1", "B2", "C1", "C2", "C3"),
  postcode = c("Me1 2re", "W1A 1AA ", "N1", " SW1A 0AA", "XX1 0XX", "sw6 1hs")
  )

data
```

<br/>

### Tidying the postcodes

As you can see, some of these postcodes are poorly formatted or incomplete. To tidy them, I'll use the `postcodes` function with a `passcode_type` parameter of `"full"`. By doing so, we exclude missing entries (such as the partial postcode of `N1`).

```{r}
data_tidy <- data |> 
  rowwise() |> 
  mutate(
    postcode_clean = postcodes(
      postcode_value = postcode, 
      postcode_type = "full"
      )
    ) |> 
  ungroup() 

data_tidy
```

I then extract the resulting postcodes into a list to use with the postcode API. (I chose not to remove `NA` values in this case but you could if many exist.)

```{r}
list_postcodes <- data_tidy |> 
  pull(postcode_clean)

list_postcodes
```

<br/>

### Getting geographic metadata on the clean postcodes

Despite this cleaning, there is still an issue, as the penultimate postcode is correctly formatted but not real. I therefore want to ensure that this postcode (`XX1 0XX`) doesn't cause our entire run to break, thereby stopping us from getting information on our final postcode. To do so, I wrap our API function in the `safely` function from the `purrr` package.

```{r}
safely_postcodes_metadata <- safely(postcodes_metadata)
```

I now call the postcode API on this list of postcodes. I then extract only the results of this call, before reshaping them.

```{r}
metadata_postcodes <- map(
  .x = list_postcodes, 
  .f = ~safely_postcodes_metadata(postcode_value = .x)
  )  

data_out <- tibble(
  postcode = list_postcodes,
  results = metadata_postcodes |> 
    map("result")
  ) |>
  tidyr::drop_na() |>
  unnest(cols = results)

data_out
```

<br/>

### Presenting our extended data

Finally, I can rejoin the results directly above with our earlier data.

```{r}
data_tidy |> 
  left_join(
    data_out, 
    by = join_by(postcode_clean == postcode)
    )
```

